Proof-Pattern Search in Coq/SSReflect

### 2015-02-04 ACMTransCL/conceptualisation.tex
Remove commented-out parts

In the previous section, we have seen how ML4PG can be used to find families of proofs following 
a common proof pattern. However, the output provided by clustering algorithms is just a set
of similar patches with no hints of why these proof-patches are deemed similar. 
% In addition, the user needs
% to replay the proof to infer the pattern. %Therefore, the conceptualisation of proof-patterns can be
% %a challenge. 
In this section, we present our approach to facilitate the understanding of proof patterns.

@@

ML4PG uses this information and produces an automaton-shape representation for discovered % to help in the visualisation of 
proof-patterns and the correlated features that determined the patterns. 
Our graphical representation is simpler than other works where automata are used to represent models that are inferred from proof traces~\cite{GWR14}.  
% This feature selection mechanism helps to explain why the proof-patches are similar since it indicates the 
% features that have been relevant for the construction of the cluster. To show the selected features and their relation with 
% the proof-patches, we introduce an automaton-shape representation. In~\cite{GWR14}, a tool to infer models 
% from proof traces was presented. Those models help to understand the behaviour exhibited by a collection
% of proofs and is represented by automata.

Generally, given a cluster of proof patches ${\cal C}$, we have an automaton $A$ with $5$ consecutive states. The $i$th state of $A$ is labelled
with the list of $i$th goals in the proof-patches contained in ${\cal C}$. The transitions between the $i$th and $i+1$th states are given by 
the $i$th tactic of each proof-patch of ${\cal C}$. If two or more tactics belong to the same group (see Figure~\ref{tab:tactics}),
they are merged in a unique transition; otherwise, the tactics will be shown as different transitions. In addition, each state is annotated with 
features 
whose correlation determined the cluster.
%for clustering that are related to that 
%state.

% $-$ there are several transitions from the $j$th to the $j+1$th state labelled by the $j$th tactic of each proof-patch of $P$ -- if either the $j$th tactic of two or more
% proof-patches is the same or belong to the same group (see Table~\ref{tab:tactics}), the transitions are merged in a unique transition labelled with the set of those tactics,
% 
% $-$ if a proof ends after the $j$th tactic of a proof-patch, the $j+1$ state is represented as final.

%\begin{example}

\begin{figure}[t]
\centering 
\includegraphics[scale=.19]{itp6.png}
\caption{\scriptsize{\emph{Fragment of the automaton 
 corresponding to the cluster of four lemma fragments described in the case study of Section~\ref{sec:recurrent}} (the whole automaton can be seen in Appendix~\ref{sec:automaton}). 
It shows correspondence between certain proof steps of lemmas % the following lemmas were discovered:
\texttt{telescope} (for the lemma of Example~\ref{example0}), \texttt{nilpotent\_inverse} (for Lemma~\ref{lem:nilpotent2}), and \texttt{persistent\_homology} 
(for Lemma~\ref{lem:fundamental}). Square boxes denote feature correlation where it exists. 
To be compact, we  hide full lemma statements, tactic and auxiliary lemma names, symbols, etc., but they can be shown by ML4PG. 
%For example, the full statements of the following cluster lemmas will be shown:
%The names of the lemmas of that cluster are: 
}}\label{fig:automata}
\end{figure}

%\end{example}

>>>>

In the previous section, we have seen how ML4PG can be used to find families of proofs following
a common proof pattern. However, the output provided by clustering algorithms is just a set
of similar patches with no hints of why these proof-patches are deemed similar.
In this section, we present our approach to facilitate the understanding of proof patterns.

@@

ML4PG uses this information and produces an automaton-shape representation for discovered
proof-patterns and the correlated features that determined the patterns.
Our graphical representation is simpler than other works where automata are used to represent models that are inferred from proof traces~\cite{GWR14}.

Generally, given a cluster of proof patches ${\cal C}$, we have an automaton $A$ with $5$ consecutive states. The $i$th state of $A$ is labelled
with the list of $i$th goals in the proof-patches contained in ${\cal C}$. The transitions between the $i$th and $i+1$th states are given by
the $i$th tactic of each proof-patch of ${\cal C}$. If two or more tactics belong to the same group (see Figure~\ref{tab:tactics}),
they are merged in a unique transition; otherwise, the tactics will be shown as different transitions. In addition, each state is annotated with
features
whose correlation determined the cluster.

\begin{figure}[t]
\centering
\includegraphics[scale=.19]{itp6.png}
\caption{\scriptsize{\emph{Fragment of the automaton
 corresponding to the cluster of four lemma fragments described in the case study of Section~\ref{sec:recurrent}} (the whole automaton can be seen in Appendix~\ref{sec:automaton}).
It shows correspondence between certain proof steps of lemmas
\texttt{telescope} (for the lemma of Example~\ref{example0}), \texttt{nilpotent\_inverse} (for Lemma~\ref{lem:nilpotent2}), and \texttt{persistent\_homology}
(for Lemma~\ref{lem:fundamental}). Square boxes denote feature correlation where it exists.
To be compact, we  hide full lemma statements, tactic and auxiliary lemma names, symbols, etc., but they can be shown by ML4PG.

}}\label{fig:automata}
\end{figure}

### 2015-02-04 ACMTransCL/conclusions.tex
Remove commented-out parts

We have presented several enhancements of the ML4PG system. \emph{Term clustering} adds a new 
functionality to ML4PG: the user can receive suggestions about families of similar definitions, types and lemma shapes (in fact, any Coq terms). 
The \emph{proof-patch method} is employed to analyse the properties of the patches that constitute a proof. The whole syntax of Coq libraries is now subject to \emph{recurrent clustering}, which %combination
%of term clustering and proof-patches
 increases the number and accuracy of families of similar proofs suggested by ML4PG. In addition, this information is taken into account in an automatic proof-generation method. Finally, the different visualisation tools implemented in ML4PG facilitates the interpretation of clusters of similar proof-patches and terms.

@@

%  \textbf{Related Work}
%  on using statistical machine-learning in ITPs was focused on speeding up proof automation in different ITPs~\cite{lpar-urban,mash,UrbanSPV08}. 
%  Comparing to these tools, we use unsupervised, rather than supervised, learning; and we do not use sparse  machine-learning methods.
%  (See also \cite{KHG13} for a detailed comparison of different machine learning tools applied in various theorem provers.)
%  We do not have a quantitative target when it comes to improving \emph{interactive} proof building experience: 
%  no longer speed up in automated proof search or the number of automatically proven theorems are the main criteria of success.
%  Instead, the user experience is the main parameter we target.
%  We generally follow the ``qualitative'' intuition that ML4PG, being an interactive hint generator, must provide interesting and non-trivial hints on user's demand,
%  and should be flexible and fast enough to do so in real  time,  at any stage of the proof, and relative to any chosen proof library.

>>>>

We have presented several enhancements of the ML4PG system. \emph{Term clustering} adds a new
functionality to ML4PG: the user can receive suggestions about families of similar definitions, types and lemma shapes (in fact, any Coq terms).
The \emph{proof-patch method} is employed to analyse the properties of the patches that constitute a proof. The whole syntax of Coq libraries is now subject to \emph{recurrent clustering}, which
 increases the number and accuracy of families of similar proofs suggested by ML4PG. In addition, this information is taken into account in an automatic proof-generation method. Finally, the different visualisation tools implemented in ML4PG facilitates the interpretation of clusters of similar proof-patches and terms.

@@

### 2015-02-04 ACMTransCL/patches.tex
Remove commented-out parts

 In~\cite{KHG13}, 
we introduced a feature extraction method for Coq proofs capturing
the user's interaction through the applied tactics. That method 
%tracks simple, 
traced 
low-level properties present in proof's subgoals, e.g. ``the
top symbol'' or ``the argument type''. Further, these features were taken
in relation to the statistics of user actions on every subgoal: how many and what
kind of tactics he applied, and what kind of arguments he provided to the
tactics. Finally, a few proof-steps were taken in relation to each other.
This method  had two drawbacks.\\ 
% (1) It only captures information 
% about the top symbol of the goal; this means that information is mainly collected 
% from symbols such as \lstinline?forall? or \lstinline?equal? (common to most proofs)
% and features from other symbols (that are likely more relevant for the proof) are never captured.
(1) It was  focused on the first
five proof-steps of a proof; therefore, some information was lost. We address this issue by implementing automatic split of each proof into proof-patches; thus, allowing ML4PG to analyse a proof by the properties
of the patches that constitute the proof. \\
(2) The method assigned most feature-values blindly; thus, being insensitive to many important parameters, such as e.g. the structure of lemmas and hypotheses used as 
tactic arguments within a proof. 
The previous section gave us the way of involving all Coq objects into recurrent feature re-evaluation.
%, and, as we have explained in the previous section, machine-learning algorithms are sensitive to 
%the assignment of feature values. These drawbacks have been overcome combining a new method called the 
%\emph{proof-patch method} and the term feature extraction method presented in the previous section.

%We start with introducing \emph{proof-patches} to address the first drawback. The underlying idea is that

%We start defining the notions of Coq proof and proof-patch matrices.

@@

From proof-patches, we can construct their feature matrices. %that will be provided to the machine-learning algorithms. 
We will shortly define the feature extraction function $[.]_P=\langle[.]_M,[.]_{tac}\rangle : proof~patches \rightarrow M[\mathbb{Q}]_{5\times 6}$,
where $[.]_{tac}$ is an injective function that has been introduced to assign values to tactics.
%To be more concrete, 
We have defined two versions of  $[.]_{tac}$: one for Coq tactics and another for SSReflect tactics. 
% -- although SSReflect is an extension of Coq, this package implements a set of proof tactics designed to support the extensive
% use of small-scale reflection in formal proofs~\cite{SSReflect}. 
In the SSReflect case, we divide the tactics into 7 groups and assign similar
values to each tactic in the group, see Table~\ref{tab:tactics}. 
% If several tactics are concatenated using ``\lstinline?;?'', the value returned by
% $[.]_{tac}$ is the concatenation of the values of the individual tactics. 
Analogously for Coq tactics, cf. Appendix~\ref{sec:coqtactics}.

%These clusters in turn are used to define a new function $[.]_{term}$. The function $[.]_{term}$ assigns numbers using the 
%formula $[l]_{term}=2\times j + p$ where $C_j$ is a cluster such that $l\in C_j$ and $p$ is the proximity value of $l$ in $C_j$. This process assigns 
%closer values to statements in the same cluster, and more distant numbers across groups. This function will be used in the following section.

@@

We finish this section with a case study that illustrates 
the use of the proof-patch method, and shows the differences with the results obtained with the old method~\cite{KHG13}. This case study concerns the discovery of proof patterns in mathematical proofs
across formalisations of apparently disjoint mathematical theories: Linear Algebra, Combinatorics and Persistent Homology  (across 758 lemmas and 5~libraries).
In this scenario, we use statistically discovered proof patterns to advance the proof of a given ``problematic'' lemma. 
Namely,  a few initial steps in its proof are clustered against several mathematical libraries. 
We deliberately take lemmas belonging to very different SSReflect libraries. The lemma introduced in Example~\ref{example0} is a basic fact about
summations. Lemma~\ref{lem:nilpotent} states a result about \emph{nilpotent} matrices %~\cite{BR91} 
(a square matrix $M$ is \emph{nilpotent}
if there exists an $n$ such that $M^n=0$). Finally, Lemma~\ref{lem:fundamental} is a generalisation of the \emph{fundamental lemma of 
Persistent Homology}~\cite{HCMS12}.

@@

 \multicolumn{7}{c}{}\\
% Second table 
\hline
 & \emph{tactics} & \emph{n} & \emph{arg type} & \emph{arg} & \emph{symbols} &  \emph{goals} \\
\hline
\hline
\emph{g1} & $([move/]_{tac},[\texttt{->}]_{tac},0,0)$ & $2$  & $([Prop]_{type},0,0,0)$ &  $([eq\_refl]_{term},0,0,0)$  & $([=]_{term},[+]_{term},[-]_{term})$ & $1$ \\
 \hline
  \emph{g2} & $([rewrite]_{tac},0,0,0)$ & $1$  & $([Prop]_{type},0,0,0)$  & $([subr\_eq0]_{term},0,0,0)$ & $([=]_{term},[+]_{term},[-]_{term})$ & $1$ \\
 \hline
%   \emph{g3} & - & -  & - & -  & - & -  \\
%  \hline
%   \emph{g4} & - & -  & - & -  & - & -  \\
%  \hline 
%   \emph{g5} & - & -  & - & -  & - & - \\
%  \hline

@@

When proving Lemma~\ref{lem:nilpotent}, the user may call ML4PG after completing a few standard proof steps: apply induction and solve the base case using rewriting.
At this point it is difficult, even for an expert user, % 60 B?
to get the intuition that he could reuse the proofs from Example~\ref{example0} and Lemma~\ref{lem:fundamental}. There are several reasons for this.
First of all, the formal proofs of these lemmas are in different libraries; then, it is difficult to establish a conceptual connection among them. Moreover,
although the three lemmas involve summations,
the types of the terms of those summations are different. Therefore, search based on types or keywords would not help. Even search  % *a* search removed -- because 59B??? 
of all the lemmas involving summations does not provide a clear suggestion, since there are more than $250$ lemmas (a considerable number for handling them manually) --- a clever search-pattern will considerably reduced the number of suggested lemmas, but such a pattern is not trivial at all.

% \begin{table}
%  	\centering
%  	\tiny{
%  		\begin{tabular}{|l|l|}
%  		\hline
%  	Goals and Subgoals & Proof-Steps (Tactics) \\
%  		\hline
%  		\hline
%  	{\scriptsize $\forall~(M:M_n) (m:nat), M^m=0 \implies (1-M)\times \sum\limits_{i=0}^{m-1} M^i = 1$} & \\
%  			& {\scriptsize \lstinline?move => M m nilpotent.?} \\
%          {\scriptsize $(1-M)\times \sum\limits_{i=0}^{m-1} M^i = 1$} & \\
%         &   {\scriptsize\lstinline?rewrite big_distrr mulmxBr mul1mx.?}\\
%         {\scriptsize$   \sum\limits_{i=0}^{m-1} M^i - M^{i+1}$} &\\
%         
%         & {\scriptsize\lstinline?case : n.?}\\
%         {\scriptsize $ \forall~(M:M_0) (m:nat),M^m=0 \implies \sum\limits_{i=0}^{m-1} M^i - M^{i+1} $ }&\\
%         &  {\scriptsize\lstinline?by rewrite !thinmx0.?}\\
%          {\scriptsize$ \forall~(M:M_{n+1}) (m:nat), M^m=0 \implies \sum\limits_{i=0}^{m-1} M^i - M^{i+1} $}&\\
%  		\hline
%  		\end{tabular}
%  		
%  		}
%  	\caption{\textbf{First steps of the proof of Lemma~\ref{lem:nilpotent},} \emph{where the user stops and invokes ML4PG for a hint.}}
%  	\label{tab:sumfirstn}
%  \end{table}

@@

%\begin{IP}\label{ip:math}
%This proof pattern is followed by Lemmas~\ref{lem:nilpotent},~\ref{lem:fundamental} and~\ref{lem:lemma3}.

@@

\emph{
%\begin{enumerate}
  %\item 
	Apply case on $n$.
  \begin{enumerate}
   \item Prove the base case (a simple task).
   \item Prove the case $0<n$:
     \begin{enumerate}
     \item expand the summation,
     \item cancel the terms pairwise,
     \item the terms remaining after the cancellation are the first and the last one. 
     \end{enumerate}
  \end{enumerate}
% \end{enumerate}
}

@@

% \begin{table}
%  	\centering
%  	\footnotesize{
%  		\begin{tabular}{|l|l|}
%  		\hline
%  	Goals and Subgoals & Proof-Steps (Tactics) \\
%  		\hline
%  		\hline
%  	{\scriptsize $\forall~(M:M_n) (m:nat), M^m=0 \implies \exists N, N \times (1-M)= 1$} & \\
%  			& {\scriptsize \lstinline?move => M m nilpotent.?} \\
%          {\scriptsize $\exists N, N \times (1-M)= 1$} & \\
%         &   {\scriptsize\lstinline?exists?}\\
%         &   {\scriptsize~~\lstinline?\sum_(0<=i<m.+1) (pot_matrix M i).?}\\
%         {\scriptsize$   (\sum\limits_{i=0}^{m-1} M^i) \times (1 - M) $} &\\
%         
%         & {\scriptsize\lstinline?rewrite big_distrl mulmxrB mulmx1.?}\\
%         {\scriptsize $ \sum\limits_{i=0}^{m-1} M^i - M^{i+1} $ }&\\
%                 & {\scriptsize\lstinline?case : n.?}\\
%         {\scriptsize $ \forall~(M:M_0) (m:nat),M^m=0 \implies \sum\limits_{i=0}^{m-1} M^i - M^{i+1} $ }&\\
%         &  {\scriptsize\lstinline?by rewrite !thinmx0.?}\\
%          {\scriptsize$ \forall~(M:M_{n+1}) (m:nat), M^m=0 \implies \sum\limits_{i=0}^{m-1} M^i - M^{i+1} $}&\\
%  		\hline
%  		\end{tabular}
%  		
%  		}
%  	\caption{\textbf{First steps of the proof of Lemma~\ref{lem:nilpotent2}.} \emph{Note the correlation between this proof and the proof of Lemma~\ref{lem:nilpotent} in Table~\ref{tab:sumfirstn}.}}
%  	\label{tab:nilpotent2}
%  \end{table}

@@

%There are two enhancements when we use the proof-patch method instead of the method presented in~\cite{KHG13}: 
The new method brings two improvements:
(1) the number of suggestions is increased, but, at the same time, (2) 
the clusters are more accurate. The proof-patch method considers fragments of proofs that are deep in the proof and were not considered before; therefore, it can find lemmas (more precisely patches of lemmas) that were not
included previously in the clusters. In our case study, ML4PG suggests two additional interesting proof fragments. The first one is an intermediate patch of the proof of 
Lemma~\ref{lem:fundamental}; then, two patches are suggested from this 
lemma: the proof-patch of the inner sum, and the proof-patch of the outer sum (both of them following Proof Strategy~\ref{ps:math}). The following lemma is also 
suggested.

@@

%The proof-patch method increases the number of suggestions; this could be a problem if more irrelevant lemmas were suggested. However, 
The new numbering of features produces more accurate clusters removing the irrelevant lemmas. In particular, using the default settings, 
ML4PG only suggests the lemma from Example~\ref{example0}, the two patches from Lemma~\ref{lem:fundamental} and the last patch from 
Lemma~\ref{lem:nilpotent2}. If the granularity is increased, the last patch from 
Lemma~\ref{lem:nilpotent2} is the only suggestion --- note that this is the closest lemma to Lemma~\ref{lem:nilpotent}.

%The discovery of proof-patterns can be a challenging task, but increasing the number and precision of suggestions simplifies this problem. 
%With the same goal (facilitate the discovery of proof-pattens), we introduce a graphical representation in the next section.

% the four lemmas given above follow the Strategy \ref{ps:math} is our next benchmarking task for ML4PG.
% Table~\ref{tab:compare0} shows the results that ML4PG will obtain, if the user varies the clustering algorithms and the size of clusters when 
% calling ML4PG within the proof environment; we also highlight the exact solution to our benchmark task, given by calling K-means algorithm with granularity $4$.

% \begin{table}
% \centering
% {\scriptsize 
%    \begin{tabular}{|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|}
%      \hline
%     & $g=1$& $g=2$& $g=3$& $g=4$ & $g=5$\\
%       Algorithm: & ($n=84$) & ($n=94$) & ($n=108$) & ($n=126$) & ($n=151$)\\     \hline
%      K-means &76$^{a-d}$ &51$^{a-d}$  &16$^{a-d}$ &\textbf{4}$^{a-d}$&2$^{b,d}$  \\
%      \hline
%      E.M. &26$^{a-d}$  &51$^{a,b,d}$ &41$^{a,b,d}$  &11$^{b,c,d}$ &3$^{a,b,d}$ \\
%      \hline
%      FarthestFirst &81$^{a-d}$ &48$^{a-d}$ &31$^{a-d}$  &25$^{a-d}$  &21$^{a-d}$  \\
%      \hline
%     \end{tabular}      }
%           
%    \caption{\textbf{Clustering experiments discovering Proof Strategy~\ref{ps:math}.} \emph{When  $g$ is chosen by the user, 
%    ML4PG dynamically calculates the number of clusters $n$. Choosing Lemma~\ref{lem:nilpotent} for pattern-search,  
%   the table shows the size of the single cluster that ML4PG displays to the user for every choice of $g$.
%    The table shows sizes of clusters containing: $a)$ Lemma~\ref{lem:nilpotent}, $b)$ Lemma~\ref{lem:lemma3},  $c)$ Lemma~\ref{lem:fundamental} and
%    $d)$ Lemma~\ref{lem:nilpotent2}. Note that with all variations of the learning algorithms and parameters, ML4PG is consistently grouping the
% correct lemmas into clusters, albeit with varied degree of precision. Note also the effect of the granularity parameter, the smallest granularity value $g=1$ produces big 
% clusters that are difficult to inspect; on the contrary, the biggest value $g=5$ produces small clusters that might not contain enough elements to spot a pattern. 
% The most accurate result for this example is obtained using the K-means algorithm and 4 as granularity parameter.}}\label{tab:compare0}
%   \end{table}
%  
% 
% Table~\ref{tab:compare0} gives an insight on how granularity parameter $g$ can be used in pattern-recognition within ML4PG.
%  One can use a ``bottom-up approach'' to detect patterns starting with the granularity value of 1 and increasing that value in successive calls.
% Using the default values $g=3$ and K-means algorithm, ML4PG obtains 15 suggestions related to lemmas about summations including Lemmas~\ref{lem:lemma3} 
% and~\ref{lem:fundamental}. Increasing the granularity level to 4, ML4PG discovers that Lemma  \ref{lem:nilpotent} is similar to 
% Lemmas~\ref{lem:lemma3}, \ref{lem:fundamental} and \ref{lem:nilpotent2}. Finally, increasing the granularity level to the maximum level of $5$, 
% ML4PG just discovers that Lemmas \ref{lem:nilpotent} and  \ref{lem:nilpotent2} are similar. %and also the following lemma;

% When the proof analogy is found (as all four proof patches taken by ML4PG follow steps $1$ and $2(a)$ of Proof Strategy \ref{ps:math}), the user can use the remaining proof steps for Lemma~\ref{lem:nilpotent2} as a template for reconstructing the remaining proof for Lemma~\ref{lem:nilpotent} (by analogy with the steps $2(b)$ and $2(c)$ in Lemma~\ref{lem:nilpotent2}). % in case of the four lemmas mentioned here, the remaining steps all follow the Strategy \ref{ps:math}. Thus, ML4PG finds patterns arising from 
% The heterogeneous cluster obtained in this scenario belongs to the category of clusters consisting of theorems that share the same proof structure and as well as an 
% auxiliary lemma (e.g. the lemma used to expand the summation).

>>>>

 In~\cite{KHG13},
we introduced a feature extraction method for Coq proofs capturing
the user's interaction through the applied tactics. That method
traced
low-level properties present in proof's subgoals, e.g. ``the
top symbol'' or ``the argument type''. Further, these features were taken
in relation to the statistics of user actions on every subgoal: how many and what
kind of tactics he applied, and what kind of arguments he provided to the
tactics. Finally, a few proof-steps were taken in relation to each other.
This method  had two drawbacks.\\
(1) It was  focused on the first
five proof-steps of a proof; therefore, some information was lost. We address this issue by implementing automatic split of each proof into proof-patches; thus, allowing ML4PG to analyse a proof by the properties
of the patches that constitute the proof. \\
(2) The method assigned most feature-values blindly; thus, being insensitive to many important parameters, such as e.g. the structure of lemmas and hypotheses used as
tactic arguments within a proof.
The previous section gave us the way of involving all Coq objects into recurrent feature re-evaluation.

@@

From proof-patches, we can construct their feature matrices.
We will shortly define the feature extraction function $[.]_P=\langle[.]_M,[.]_{tac}\rangle : proof~patches \rightarrow M[\mathbb{Q}]_{5\times 6}$,
where $[.]_{tac}$ is an injective function that has been introduced to assign values to tactics.
We have defined two versions of  $[.]_{tac}$: one for Coq tactics and another for SSReflect tactics.
In the SSReflect case, we divide the tactics into 7 groups and assign similar
values to each tactic in the group, see Table~\ref{tab:tactics}.
Analogously for Coq tactics, cf. Appendix~\ref{sec:coqtactics}.

@@

We finish this section with a case study that illustrates
the use of the proof-patch method, and shows the differences with the results obtained with the old method~\cite{KHG13}. This case study concerns the discovery of proof patterns in mathematical proofs
across formalisations of apparently disjoint mathematical theories: Linear Algebra, Combinatorics and Persistent Homology  (across 758 lemmas and 5~libraries).
In this scenario, we use statistically discovered proof patterns to advance the proof of a given ``problematic'' lemma.
Namely,  a few initial steps in its proof are clustered against several mathematical libraries.
We deliberately take lemmas belonging to very different SSReflect libraries. The lemma introduced in Example~\ref{example0} is a basic fact about
summations. Lemma~\ref{lem:nilpotent} states a result about \emph{nilpotent} matrices
(a square matrix $M$ is \emph{nilpotent}
if there exists an $n$ such that $M^n=0$). Finally, Lemma~\ref{lem:fundamental} is a generalisation of the \emph{fundamental lemma of
Persistent Homology}~\cite{HCMS12}.

@@

 \multicolumn{7}{c}{}\\
% Second table
\hline
 & \emph{tactics} & \emph{n} & \emph{arg type} & \emph{arg} & \emph{symbols} &  \emph{goals} \\
\hline
\hline
\emph{g1} & $([move/]_{tac},[\texttt{->}]_{tac},0,0)$ & $2$  & $([Prop]_{type},0,0,0)$ &  $([eq\_refl]_{term},0,0,0)$  & $([=]_{term},[+]_{term},[-]_{term})$ & $1$ \\
 \hline
  \emph{g2} & $([rewrite]_{tac},0,0,0)$ & $1$  & $([Prop]_{type},0,0,0)$  & $([subr\_eq0]_{term},0,0,0)$ & $([=]_{term},[+]_{term},[-]_{term})$ & $1$ \\
 \hline

@@

When proving Lemma~\ref{lem:nilpotent}, the user may call ML4PG after completing a few standard proof steps: apply induction and solve the base case using rewriting.
At this point it is difficult, even for an expert user,
to get the intuition that he could reuse the proofs from Example~\ref{example0} and Lemma~\ref{lem:fundamental}. There are several reasons for this.
First of all, the formal proofs of these lemmas are in different libraries; then, it is difficult to establish a conceptual connection among them. Moreover,
although the three lemmas involve summations,
the types of the terms of those summations are different. Therefore, search based on types or keywords would not help. Even search
of all the lemmas involving summations does not provide a clear suggestion, since there are more than $250$ lemmas (a considerable number for handling them manually) --- a clever search-pattern will considerably reduced the number of suggested lemmas, but such a pattern is not trivial at all.

@@

@@

\emph{
	Apply case on $n$.
  \begin{enumerate}
   \item Prove the base case (a simple task).
   \item Prove the case $0<n$:
     \begin{enumerate}
     \item expand the summation,
     \item cancel the terms pairwise,
     \item the terms remaining after the cancellation are the first and the last one.
     \end{enumerate}
  \end{enumerate}
}

@@

@@

The new method brings two improvements:
(1) the number of suggestions is increased, but, at the same time, (2)
the clusters are more accurate. The proof-patch method considers fragments of proofs that are deep in the proof and were not considered before; therefore, it can find lemmas (more precisely patches of lemmas) that were not
included previously in the clusters. In our case study, ML4PG suggests two additional interesting proof fragments. The first one is an intermediate patch of the proof of
Lemma~\ref{lem:fundamental}; then, two patches are suggested from this
lemma: the proof-patch of the inner sum, and the proof-patch of the outer sum (both of them following Proof Strategy~\ref{ps:math}). The following lemma is also
suggested.

@@

The new numbering of features produces more accurate clusters removing the irrelevant lemmas. In particular, using the default settings,
ML4PG only suggests the lemma from Example~\ref{example0}, the two patches from Lemma~\ref{lem:fundamental} and the last patch from
Lemma~\ref{lem:nilpotent2}. If the granularity is increased, the last patch from
Lemma~\ref{lem:nilpotent2} is the only suggestion --- note that this is the closest lemma to Lemma~\ref{lem:nilpotent}.
